---
title: "Database of borrowed lexicon in Daghestan (pilot)"

output: 
    html_document: 
    theme: spacelab
    highlight: tango 
    toc: yes 
    toc_float: yes
    smooth_scroll: false
    fig_cap: TRUE

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

This *try-out* database contains wordlists collected in the field in Daghestan, as part of the [Daghestanian loans](https://ilcl.hse.ru/en/projects) project by the [Linguistic Convergence Laboratory](https://ilcl.hse.ru/en/) at NRU HSE. The aim of the 160-item shortlist, which is based on the [World Loanword Database questionnaire](https://wold.clld.org/), is to measure lexical contact on a micro-level. In other words, to quantify lexical convergence among the speech communities of minority languages on a village-level, and to detect fine-grained areal patterns beyond general observations on the spheres of influence of certain languages.


```{r, echo=FALSE, message=FALSE, warning=FALSE}

# loading packages

library(tidyverse)
library(lingtypology)
library(DT)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}


# Load data on recorded lists

speaker_data <- read_tsv("speakers_villages.csv")


# And on dictionary data

dict <- read_tsv("dictionaries.csv")


# Create a dataframe that counts the number of lists for each village

lists <- speaker_data %>%
  group_by(Village, Language, Latitude, Longitude) %>%
  summarise(Lists = n())


# Convert the column with these counts to a factor

lists$Lists <- as.factor(lists$Lists)


```


## Map of the surveyed villages

The map below shows the villages surveyed and the number of lists collected. White dots represent dictionary data. Hover over a dot to see the language it represents. When you click on a dot, a popup will appear that shows genetic info on the language, the name of the village or in case of the white dots, the resource used.

<br>

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 9, fig.height= 7}


# Draw a map that shows each village + the numer of lists collected


map.feature(lists$Language,
            feature = lists$Lists,
            latitude = lists$Latitude,
            longitude = lists$Longitude,
            zoom.control = TRUE,
            legend = TRUE,
            title = "Lists",
            label = lists$Language,
            popup = paste(aff.lang(lists$Language), # Popup with genetic info on language + name of village
                          "<br>",
                          "<br>",
                          "<b> Village: </b>",
                          lists$Village),
            zoom.level = 7.5) %>% # Put another map on top of it, showing dictionary data
              map.feature(dict$Language,
              latitude = dict$Latitude,
              longitude = dict$Longitude,
              feature = dict$Language,
              label = dict$Language,
              popup = paste(aff.lang(dict$Language),
                            "<br>",
                            "<br>",
                            "<b>Source:</b>",
                            dict$Source),
              legend = FALSE,
              color = c("white"),
              tile = c("Esri.WorldGrayCanvas"),
              pipe.data = .)
```

<br>

## The list

Below is the shortlist of concepts and the Russian translation used to elicit in the field.

<br>

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Load data

zilo <- read_tsv("zilo.csv")

core <- zilo[!duplicated(zilo$`Source concept`),]
core <- core[-c(1:3, 9:19)]


datatable(core)


```


<br>

### List length

While the list of concepts consists of 160 items, the collected lists rarely do. There can be two or more translations for a single concept, or speakers might not remember how to translate all of the concepts. Below is a density plot showing the dispersion of wordlist length.


```{r, echo=FALSE, message=FALSE, warning=FALSE}


# Throw out data from dictionaries

nondict <- zilo[zilo$Language == 'Andi',]

# Count number of items recorded for each respondent

length <- nondict %>%
  group_by(Respondent) %>%
  summarise(`Target word` = n())


# Add the length of the original shortlist

shortlist <- c("shortlist", 160)
length <- rbind(length, shortlist)
length$`Target word` <- as.numeric(length$`Target word`)


# Create a density plot

plot(density(length$`Target word`), main = "Dispersion of wordlist length", xlab = "Length")


```

<br>

### Number of loans in the sample by donor language

<br>

```{r, echo=FALSE, message=FALSE, warning=FALSE}

loans <- nondict %>%
  group_by(Respondent, `Supposed toplevel source`) %>%
  summarise(`Target word` = n())

ggplot(loans, aes(x = `Supposed toplevel source`, 
                 y = `Target word`)) +geom_bar(stat="identity", 
                                            aes(fill = `Supposed toplevel source`)) + labs(x = "", y = "", fill = "Donor language")

```


<br>

---

## The Database

<br>


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10}

# preparing data

zilo <- read_tsv("zilo.csv")

# Create column for rownames

zilo$Entry <- paste(zilo$`Entry number`, zilo$`Target word`, sep = " ")


# Drop some columns I don't want to show for now

zilo_database <- zilo[-c(1:3, 6, 13:14, 16:20)]
row.names(zilo_database) <- zilo$Entry

```

```{r, echo=FALSE, message=FALSE}

# Create a searchable datatable with filters on top

datatable(zilo_database, class = 'cell-border stripe', filter = 'top', options = list(pageLength = 50))


```

<br>

---

Version: `r Sys.Date()`. For questions or comments contact jh.verhees@gmail.com.

---

